---
title: "data-finalproj"
author: "Feven Yohannes"
date: "November 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(readr)
library(dplyr)
library(knitr)
library(ggplot2)
library(stringr)
library(janitor)
library(gt)
library(httr)
library(rvest)
library(dplyr)
library(fs)
library(sf)
library(gganimate)
library(tidyverse)
library(leaflet)
library(gganimate)
```

Clean data for the UN
```{r}
un <- read_html("http://data.un.org/en/iso/et.html", skip=4, col_names=TRUE)
tbls <- html_nodes(un, "table")
un_tbl <- un %>%
        html_nodes("table") %>%
        .[3:4] %>%
        html_table(fill = TRUE, header=TRUE)

un_test <- un_tbl %>% map_df(bind_rows) #Merged two tibbles within one tibble
  
colnames(un_test)[1] <- "indicators" #Renamed first column
colnames(un_test)[5] <- "indicators2" #Renamed fifth column


un_test<- un_test  %>%
  unite(col="merged",c("indicators","indicators2"),sep="", na.rm=TRUE)
#Merged both of the indicators together, so it could fall under one column.
```

# Reshaping data for graph purposes
```{r}
un_graph <- un_test %>%
  pivot_longer(cols=c(`2005`,`2010`,`2018`),names_to= "Years") 
  
```

#Removing extra characters in values 
```{r}
un_graph <- un_graph %>%
  mutate(value_new= (str_remove_all(value, "[bdlc]"))) 
``` 

#Writing data into an Rds file
```{r}
write_rds(un_graph,"Ethiopia_Economic_Development/UN.rds")
```


#First graphy for shiny demo
```{r}
un_graph %>%
  filter(str_detect(merged,fixed("Economy: Agriculture"))) %>%
  ggplot(aes(x=Years,y=value,group=merged)) + 
  geom_point() +
  geom_line()
```

#Growth of Businesses Graph 
#Note: Separate values for males and females 
```{r}
un_graph %>%
  filter(str_detect(merged,fixed("Labour force participation ratee"))) %>%
  ggplot(aes(x=Years,y=value_new,group=merged)) + 
  geom_point() +
  geom_line()
```

UN Graph Life Expectancy 
```{r}
un_graph %>%
  filter(str_detect(merged,fixed("Life expectancy at birthj"))) %>%
  filter(str_detect(merged,fixed("male"))) %>%
  ggplot(aes(x=Years,y=value_new,group=merged)) + 
  geom_point() +
  geom_line()

```


#UN Economic graph, FACET GRID FOR JOBS
```{r}
un_graph %>%
 filter(str_detect(merged,fixed("Economy"))) %>%
  ggplot(aes(x=Years,y=value_new)) +geom_point(color="green") +
  #Changing the alpha transparency of the points
  geom_smooth(se = FALSE, method = "lm") +
  #Adding a best fit line
  facet_grid(. ~ merged) +
  labs(title = "Change in job sectors",
              x = "Years",
              y= "% of employed population")

#String split for merged indicator and units
#Strip the parathenses 

   

```

#Clean data for the World bank
```{r}
world_bank <- read.csv("/Users/feven/Downloads/API_ETH_DS2_en_csv_v2_205341 (5).csv", header = TRUE, sep = ",", na.strings = ".") 

```







Re-formating the world bank data so all the years are in one column. Makes it easier to graph.
```{r}
world_bank <- world_bank %>%
  pivot_longer(cols=c(`X1960`,`X1961`,`X1962`,`X1963`,`X1964`,`X1965`,`X1966`,`X1967`,`X1968`,`X1969`,`X1970`,`X1971`,`X1972`,`X1973`,`X1974`,`X1975`,`X1976`,`X1977`,`X1978`,`X1979`,`X1980`, `X1981`,`X1982`,`X1983`,`X1984`,`X1985`,`X1986`,`X1987`,`X1988`,`X1989`,`X1990`,`X1991`,`X1992`,`X1993`,`X1994`,`X1995`,`X1996`,`X1997`,`X1998`,`X1999`,`X2000`,`X2001`,`X2002`,`X2003`,`X2004`,`X2005`,`X2006`,`X2007`,`X2008`,`X2009`,`X2010`,`X2011`,`X2012`,`X2013`,`X2014`,`X2015`,`X2016`,`X2017`,`X2018`),names_to= "Years") 
```

Mutating for Years without the x's in front of it
```{r}
world_bank <- world_bank %>% mutate(Years_new = as.numeric(substr(Years,2,5)))
```

Dropping NA's for the world bank
```{r}
world_bank$value <- as.character(world_bank$value)

world_bank %>%
  drop_na()
```


```{r}
world_bank <- world_bank %>%
      filter(value != "NA" )
 
world_bank$value <- as.numeric(world_bank$value)
```

Reading World Bank Data into rds
```{r}
write_rds(world_bank,"Ethiopia_Economic_Development/Worldbank.rds")
```




World Bank, mortality rate for infant graphs
```{r}
world_bank %>%
  filter(str_detect(Indicator.Name,fixed("Mortality rate, under-5"))) %>%
  ggplot(aes(x=Years_new,y=value)) +geom_point() + scale_x_continuous(breaks = seq(1970,2018,by = 5)) +scale_x_continuous(breaks = seq(0,244.7,by = 10))   
  #I want to see it in general, not for each gender
```

World Bank, Incidence of Malaria
```{r}
world_bank %>%
  filter(str_detect(Indicator.Name,fixed("Incidence of malaria"))) %>%
  ggplot(aes(x=Years_new,y=value)) +geom_point()
    
```

World Bank, Maternal Deaths
```{r}
world_bank %>%
  filter(str_detect(Indicator.Name,fixed("Number of maternal deaths"))) %>%
  ggplot(aes(x=Years_new,y=value)) + geom_point()
```

World Bank, Primary School Enrollment by Gender and Total
```{r}
world_bank %>%
  filter(str_detect(Indicator.Name,fixed("School enrollment, primary"))) %>%
  filter(!str_detect(Indicator.Name,"private")) %>%
  filter(str_detect(Indicator.Name,fixed("% gross"))) %>%
    ggplot(aes(x=Years_new,y=value,color=Indicator.Name)) +geom_point() +
  geom_smooth(se = FALSE, method = "lm")
```

World Bank, 
```{r}
world_bank %>%
  filter(str_detect(Indicator.Name,fixed("Access to electricity (% of population)"))) %>%
    ggplot(aes(x=Years_new,y=value)) +geom_point() 
```




World Bank, C02 emissons
```{r}
world_bank %>%
  filter(str_detect(Indicator.Name,fixed("CO2 emissions (metric tons per capita)")))
```

World Bank, Electricity Access
```{r}
world_bank %>%
  filter(Indicator.Name == ("Access to electricity(% of population)"))
```
World Bank, Urban Population 
```{r}
world_bank %>%
  filter(str_detect(Indicator.Name,fixed("Urban population (% of total population)")))
```
World Bank, Total Population #maybe combine both graphs
```{r}
world_bank %>%
  filter(str_detect(Indicator.Name,fixed("Population, total")))
```

World Bank, Taxes on Income 
Taxes on income, profits and capital gains
```{r}
world_bank %>%
  filter(str_detect(Indicator.Name,fixed("Taxes on income, profits and capital gains (% of revenue)"))) %>%
    ggplot(aes(x=Years_new,y=value)) +geom_point() 

```



UN, GDP
```{r}
un_graph %>%
  filter(str_detect(merged,fixed("GDP: Gross domestic product"))) %>%
  ggplot(aes(x=Years,y=value)) +geom_point()
  
```
UN, GDP per capita
```{r}
gdp<-un_graph %>%
  filter(str_detect(merged,fixed("GDP per capita"))) %>%
  ggplot(aes(x=Years,y=value)) +geom_point()
```

Dataset for Map
```{r,message=FALSE,warning=FALSE}
map <- read_csv("Ethiopia_Economic_Development/et.csv")
```

Reading the Map Data as a rds
```{r}
write_rds(map,"Ethiopia_Economic_Development/map.rds")
```


```{r}
locations <-st_as_sf(map,coords=c('lng','lat')) 

map <- ggplot() +
   geom_sf(data = map) +
   geom_sf(data = locations,aes(color= city)) +
   transition_manual(city)
  #Made it so the data was grouped by locations and looked at the different races 
   labs(title="Map of Ethiopia",
        subtitle="Looking at Different Cities in Ethiopia") 
#I gave the animation a title
   

dir_create("graphics")
#I created a directory called graphics
write_rds(map,"map.rds")
map
```
```{r}
map %>%
ggplot(aes(x=lat,y=lng)) +
    geom_sf() +
    xlab("Longitude") + ylab("Latitude")

+
    ggtitle("World map", subtitle = paste0("(", length(unique(world$NAME)), " countries)"))
```

Map of Ethiopia with Leaflet
```{r}

m <- leaflet(options = leafletOptions(dragging = FALSE,
minZoom = 5,
maxZoom = 8)) %>%
  addProviderTiles(providers$Esri.NatGeoWorldMap)  %>% 
  addMarkers(lng=map$lng, lat=map$lat, popup=map$city) 
m
```

